### 2023-11-06 paper/analysis.tex
individual files

 
% -- aginst Rich et al. BDSD survey ugriz podaci
% -- Bailer-Jones i S82
% -- Trilegal simulacije s SDSS SED modelima (validacija)
% -- kuglasti skupovi