### 2023-11-06 paper/discussion.tex

% -- popraviti modele
% -- dovući model s mladim zvijezdama
% -- dodati modele za WD, BHB, unresolved binaries, RRL i druge populacije

% Here we provide a layout of our method that simultaneously provides distance estimates as well as the three-dimensional estimation of the dust distribution. The distance modulus of a star can be estimated given its absolute and apparent magnitudes and the extinction along the line of sight. Therefore, given the observed colors it is necessary to simultaneously estimate the effective temperature, metallicity and extinction.