### 2024-01-08 paper/paper.tex
changed file names

\documentclass{aastex63} 
% \documentclass[12pt,preprint]{aastex}
% \documentclass[linenumbers, twocolumn, trackchanges]{aastex631}

\usepackage{amssymb,amsmath}
% \usepackage{lsst}
\usepackage{xspace}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{comment}
\usepackage{booktabs}
\addto\extrasenglish{%
  \renewcommand{\equationautorefname}{eq.}%
}

%%%%%%%%%%%%%%%%%%%%%%
\received{\today}
\revised{XXX}
\accepted{XXX}
\submitjournal{AJ}

% Croatian letter "dj"
\def\d   {{d $\mkern-14.3mu \mathchar'26 $}}

\usepackage{xspace}
\usepackage{amsmath}
\usepackage[T1]{fontenc}

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand{\magcom}[1]{\textcolor{magenta}{#1}} % {\textbf{#1}}}
\newcommand{\redcom}[1]{\textcolor{red}{#1}} % {\textbf{#1}}}
\newcommand{\yelcom}[1]{\colorbox{yellow}{#1}} % {\textbf{#1}}}
\newcommand{\pd}{\texttt{photoD}\xspace}
\newcommand{\mh}{\ensuremath{[M/H]}\xspace}
\newcommand{\feh}{\ensuremath{[Fe/H]}\xspace}
\newcommand{\Mb}{\ensuremath{M{\mathrm{_b}}}\xspace}
\newcommand{\Mr}{\ensuremath{M{\mathrm{_r}}}\xspace}
\newcommand{\Ar}{\ensuremath{A{\mathrm{_r}}}\xspace}
\newcommand{\EBV}{\ensuremath{E \left( B \! - \! V \right)}\xspace}
\newcommand{\logg}{\ensuremath{log(g)}\xspace}
\newcommand{\teff}{\ensuremath{T\mathrm{_{eff}}}\xspace}
\newcommand{\RV}{\ensuremath{R\mathrm{V}}\xspace}
\newcommand{\AV}{\ensuremath{A{\mathrm{_V}}}\xspace}

\newcommand{\ul}{\ensuremath{u}\xspace}
\newcommand{\gl}{\ensuremath{g}\xspace}
\newcommand{\rl}{\ensuremath{r}\xspace}
\newcommand{\il}{\ensuremath{i}\xspace}
\newcommand{\zl}{\ensuremath{z}\xspace}
\newcommand{\yl}{\ensuremath{y}\xspace}

\renewcommand{\arg}[1]{\! \left( #1 \right)}
\newcommand{\Pcond}[2]{P \arg{ #1 \, | \, #2 }}
% \newcommand{\yps}{\ensuremath{y_{\mathrm{P1}}}\xspace}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{March 1, 2021}
%\revised{April 1, 2021}
%\accepted{\today}

%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the argument.
%\submitjournal{PSJ}

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{PhotoD with LSST}
\shortauthors{Palaversa et al.}
\graphicspath{{./}{figures/}}

\begin{document}

\title{PhotoD with LSST: stellar photometric distances out to the edge of the Galaxy}

\author[0000-0003-3710-0331]{Lovro Palaversa}
\affiliation{Ru{\dj}er Bo\v{s}kovi\'{c} Institute, Bijeni\v{c}ka cesta 54, 10000 Zagreb, Croatia }

\author[0000-0012-3245-123X]{\v{Z}eljko Ivezi\'{c}}
\affiliation{Department of Astronomy and DiRAC Institute, University of Washington, Box 351580, Seattle, WA 98195, USA}

\author[0000-0012-3245-123X]{Karlo Mrakov\v{c}i\'{c}}
\affiliation{Faculty of Physics, University of Rijeka, Radmile Matej\v{c}i\'{c} 2, 51000 Rijeka, Croatia}

\author[0000-0000-0000-0000]{Bob Abel}
\affiliation{The Paradise Island, WA, USA }

\author[0000-0000-0000-0000]{Dani Chao}
\affiliation{Ru{\dj}er Bo\v{s}kovi\'{c} Institute, Bijeni\v{c}ka cesta 54, 10000 Zagreb, Croatia }

\author[0000-0002-3045-0446]{Oleksandra Razim}
\affiliation{Ru{\dj}er Bo\v{s}kovi\'{c} Institute, Bijeni\v{c}ka cesta 54, 10000 Zagreb, Croatia }

\author[0000-0000-0000-0000]{Filip Matkovi\'{c}}
\affiliation{Hvar Observatory, 10000 Zagreb, Croatia}

\author[0000-0000-0000-0000]{Connor Yablonski}
\affiliation{Department of Astronomy, University of Washington, Box 351580, Seattle, WA 98195, USA}

\author[0000-0000-0000-0000]{Toni \v{S}ari\'{c}}
\affiliation{University of Split, Croatia}

\author[0000-0000-0000-0000]{Tomislav Jurki\'{c}}
\affiliation{University of Rijeka, Croatia}

\author[0000-0000-0000-0000]{Mario Juri\'{c}}
\affiliation{Department of Astronomy and DiRAC Institute, University of Washington, Box 351580, Seattle, WA 98195, USA}

\author[0000-0000-0000-0000]{Forgotten Collaborator}
\affiliation{Somewhere}

%\correspondingauthor{Lovro Palaversa}
%\email{lp@irb.hr}

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
As demonstrated with SDSS, Pan-STARRS and most recently with Gaia data, broad-band near-UV to near-IR photometry can be used 
to estimate distances, metallicity and the interstellar dust extinction along the line of sight for stars in the Galaxy. Anticipating
photometric catalogs with tens of billions of stars from Rubin's Legacy Survey of Space and Time (LSST), we present a fully Bayesian procedure 
and pipeline that build on previous work and can handle LSST-sized datasets. Likelihood computations utilize MIST/Dartmouth isochrones
and priors are derived from TRILEGAL-based simulated LSST catalogs from Dal Tio et al. (2022). The computation speed is about 10 mas per
star for both optimized grid search and MCMC methods; we show in a companion paper by Mrakov\v{c}i\'{c} et al. how to utilize neural
networks to accelerate this performance by an order of magnitude).  We validate our pipeline, named
PhotoD (in analogy with photo-z, photometric redshifts of galaxies) using both simulated catalogs, and SDSS, DECam, and Gaia data. 
We intend to make LSST-based value-added PhotoD catalogs publicly
available via Rubin Science Platform with every LSST Data Release. 
\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
%% https://astrothesaurus.org
%% You will be asked to selected these concepts during the submission process
%% but this old "keyword" functionality is maintained in case authors want
%% to include these concepts in their preprints.
\keywords{Distance measure (395) --- Interstellar extinction (841) --- Photometry (1234) --- Stellar distance (1595) --- Two-color diagrams (1724)}

\include{intro}

\include{methodology}

%\include{methodologyZI}

%\include{analysis}

%\include{discussion}

\acknowledgements
\v{Z}.I. acknowledges funding by the Fulbright Foundation and thanks the Ru\d er Bo\v{s}kovi\'{c} Institute for hospitality.

Funding for the SDSS and SDSS-II has been provided by the Alfred
P. Sloan Foundation, the Participating Institutions, the National
Science Foundation, the U.S. Department of Energy, the National
Aeronautics and Space Administration, the Japanese Monbukagakusho, the
Max Planck Society, and the Higher Education Funding Council for
England. The SDSS Web Site is \url{https://www.sdss.org/}.

This work has made use of data from the European Space Agency (ESA)
mission Gaia \url{https://www.cosmos.esa.int/gaia}, processed by the Gaia
Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/ dpac/consortium}). Funding for the
DPAC has been provided by national institutions, in particular, the
institutions participating in the Gaia Multilateral Agreement.

\vspace{5mm}
\facilities{Sloan Digital Sky Survey, Gaia}
\software{numpy \citep{numpy}, matplotlib \citep{matplotlib}, seaborn \citep{seaborn}, scipy \citep{scipy}, pandas \cite{pandas}, 
                astropy \citep{astropy-1, astropy-2}, astroML \citep{2012cidu.conf...47V}, Jupyter \citep{jupyter}, Python \citep{python}.}

\appendix
%\include{appendix}

\bibliography{paper}{}
\bibliographystyle{aasjournal}

\end{document}

	paper/Makefile
	paper/aastex.cls
	paper/aastex63.cls
	paper/apj.bst
	paper/appendix.tex
	paper/deluxetable.sty
	paper/lsst.sty
	paper/methodologyZI.tex
	paper/paper.bib
	paper/paper.pdf
	paper/paper.tex
	paper/results.tex

        plots/2Dmap.png
	plots/2Dmap_Stripe82_MrSDSS_MrGaia.png
	plots/2Dmap_Stripe82_MrSDSS_MrGaia_piSNR20.png
	plots/2Dmap_Stripe82_MrdMr_piSNR20.png
	plots/QrCornerPlot.png
	plots/bayesPanels.png
	plots/cornerPlot3.png
	plots/margPosteriors.png
	plots/margPosteriors3D.png
	plots/qpB_Ar.png
	plots/qpB_FeH.png
	plots/qpB_Mr.png
	plots/qpB_chiTest4_3D_Ar.png
	plots/qpB_chiTest4_3D_FeH.png
	plots/qpB_chiTest4_3D_Mr.png
	plots/qpB_chiTest4_FeH.png
	plots/qpB_chiTest4_Mr.png
	plots/qpBcmd.png
	plots/qpBcmd_chiTest4_3D_Ar.png
	plots/qpBcmd_chiTest4_3D_FeH.png
	plots/qpBcmd_chiTest4_3D_Mr.png
	plots/qpBcmd_chiTest4_FeH.png
	plots/qpBcmd_chiTest4_Mr.png
	plots/qpBmeans_chiTest4_.png
	plots/qpBmeans_chiTest4_3D.png