### 2024-05-07 paper/pipeline.tex
for Lovro...

\section{All-sky PhotoD Pipeline Implementation \label{sec:LSDBpipe}}

{\bf Lovro:} will here describe LSDB implementation (with Neven's help) and pipeline performance on the full TRILEGAL catalog...

LSDB = Large Survey Database

\subsection{LSDB and Healpix Implementation}

Implementation of the Bayesian algorithm described in preceding sections assumes that priors can be assumed constant
over some sky region, hereafter a patch (see \S\ref{sec:priors}). In practice,  patches need to be at least several sq. deg.
large in order to have enough stars to compute priors. At the same time, they should not be too large to satisfy the assumption
of constancy. Both of these size limits vary across the sky due to Galactic structure effects.

One of recent tools developed to efficiently loop across the sky patches and perform some per-patch calculations
is the  Large Survey Database...