### 2024-09-01 paper/paper.tex
Final edits to the TeX.

\documentclass{aastex631} 
% \documentclass[12pt,preprint]{aastex}
% \documentclass[linenumbers, twocolumn, trackchanges]{aastex631}

\usepackage{amssymb,amsmath}
% \usepackage{lsst}
\usepackage{xspace}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{comment}
\usepackage{booktabs}
\usepackage{gensymb}

\addto\extrasenglish{%
  \renewcommand{\equationautorefname}{eq.}%
}

%%%%%%%%%%%%%%%%%%%%%%
\received{\today}
\revised{XXX}
\accepted{XXX}
\submitjournal{AJ}

% Croatian letter "dj"
\def\dj   {{d $\mkern-14.3mu \mathchar'26 $}}

\usepackage{xspace}
\usepackage{amsmath}
\usepackage[T1]{fontenc}

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand{\magcom}[1]{\textcolor{magenta}{#1}} % {\textbf{#1}}}
\newcommand{\redcom}[1]{\textcolor{red}{#1}} % {\textbf{#1}}}
\newcommand{\yelcom}[1]{\colorbox{yellow}{#1}} % {\textbf{#1}}}
\newcommand{\pd}{\texttt{photoD}\xspace}
\newcommand{\mh}{\ensuremath{[M/H]}\xspace}
\newcommand{\feh}{\ensuremath{[Fe/H]}\xspace}
\newcommand{\Mb}{\ensuremath{M{\mathrm{_b}}}\xspace}
\newcommand{\Mr}{\ensuremath{M{\mathrm{_r}}}\xspace}
\newcommand{\Ar}{\ensuremath{A{\mathrm{_r}}}\xspace}
\newcommand{\EBV}{\ensuremath{E \left( B \! - \! V \right)}\xspace}
\newcommand{\logg}{\ensuremath{log(g)}\xspace}
\newcommand{\teff}{\ensuremath{T\mathrm{_{eff}}}\xspace}
\newcommand{\RV}{\ensuremath{R\mathrm{V}}\xspace}
\newcommand{\AV}{\ensuremath{A{\mathrm{_V}}}\xspace}
\newcommand\mean[1]{\overline{#1}}

\newcommand{\ul}{\ensuremath{u}\xspace}
\newcommand{\gl}{\ensuremath{g}\xspace}
\newcommand{\rl}{\ensuremath{r}\xspace}
\newcommand{\il}{\ensuremath{i}\xspace}
\newcommand{\zl}{\ensuremath{z}\xspace}
\newcommand{\yl}{\ensuremath{y}\xspace}

\renewcommand{\arg}[1]{\! \left( #1 \right)}
\newcommand{\Pcond}[2]{P \arg{ #1 \, | \, #2 }}
% \newcommand{\yps}{\ensuremath{y_{\mathrm{P1}}}\xspace}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{March 1, 2021}
%\revised{April 1, 2021}
%\accepted{\today}

%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the argument.
%\submitjournal{PSJ}

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{PhotoD with LSST}
\shortauthors{Palaversa et al.}
\graphicspath{{./}{figures/}}

\begin{document}

\title{Photo-D with LSST: stellar photometric distances out to the edge of the Galaxy}

\author[0000-0003-3710-0331]{Lovro Palaversa}
\affiliation{Ru{\dj}er Bo\v{s}kovi\'{c} Institute, Bijeni\v{c}ka cesta 54, 10000 Zagreb, Croatia }

\author[0000-0001-5250-2633]{\v{Z}eljko Ivezi\'{c}}
\affiliation{DIRAC Institute and the Department of Astronomy, University of Washington, 3910 15th Avenue NE, Seattle, WA 98195, USA}

\author[0000-0003-3287-5250]{Neven \v{C}aplar}
\affiliation{DIRAC Institute and the Department of Astronomy, University of Washington, 3910 15th Avenue NE, Seattle, WA 98195, USA}

\author[0009-0009-8154-3827]{Karlo Mrakov\v{c}i\'{c}}
\affiliation{Faculty of Physics, University of Rijeka, Radmile Matej\v{c}i\'{c} 2, 51000 Rijeka, Croatia}

\author[0000-0001-8418-3083]{Bob Abel}
\affiliation{Olympic College, 1600 Chester Avenue, Bremerton, WA
98337, USA, and DiRAC Institute, University of Washington, Box 351580, Seattle, WA 98195, USA}

% no confirmation of authorship 
%\author[0000-0000-0000-0000]{Dani Chao}
%\affiliation{Ru{\dj}er Bo\v{s}kovi\'{c} Institute, Bijeni\v{c}ka cesta 54, 10000 Zagreb, Croatia }

\author[0000-0002-3045-0446]{Oleksandra Razim}
\affiliation{Ru{\dj}er Bo\v{s}kovi\'{c} Institute, Bijeni\v{c}ka cesta 54, 10000 Zagreb, Croatia }

\author[0009-0002-5858-585X]{Filip Matkovi\'{c}}
\affiliation{Hvar Observatory, Faculty of Geodesy, University of Zagreb, Ka\v{c}i\'{c}eva 26, 10000 Zagreb, Croatia}

\author[0009-0000-0990-8339]{Connor Yablonski}
\affiliation{DIRAC Institute and the Department of Astronomy, University of Washington, 3910 15th Avenue NE, Seattle, WA 98195, USA}

\author[0000-0001-8731-8369]{Toni \v{S}ari\'{c}}
\affiliation{University of Split - FESB, R. Bo\v{s}kovi\'{c}a 32, 21000 Split, Croatia}

\author[0000-0002-4993-2939]{Tomislav Jurki\'{c}}
\affiliation{Faculty of Physics, University of Rijeka, Radmile Matej\v{c}i\'{c} 2, 51000 Rijeka, Croatia}

\author[0000-0003-1996-9252]{Mario Juri\'{c}}
\affiliation{DIRAC Institute and the Department of Astronomy, University of Washington, 3910 15th Avenue NE, Seattle, WA 98195, USA}

%\author[0000-0000-0000-0000]{Forgotten Collaborator}
%\affiliation{Somewhere}

%\correspondingauthor{Lovro Palaversa}
%\email{lp@irb.hr}

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
As demonstrated with SDSS, Pan-STARRS and most recently with Gaia data, broad-band near-UV to near-IR stellar photometry can be used 
to estimate distance, metallicity and interstellar dust extinction along the line of sight for stars in the Galaxy. Anticipating
photometric catalogs with tens of billions of stars from Rubin's Legacy Survey of Space and Time (LSST), we present a Bayesian procedure 
and pipeline that build on previous work and can handle LSST-sized datasets. Likelihood computations utilize MIST/Dartmouth isochrones
and priors are derived from TRILEGAL-based simulated LSST catalogs from Dal Tio et al. (2022). The computation speed is about 10 milliseconds per
star on a single core for both optimized grid search and MCMC methods; we show in a companion paper by Mrakov\v{c}i\'{c} et al. how to utilize neural
networks to accelerate this performance by up to an order of magnitude.  We validate our pipeline, named
PhotoD (in analogy with photo-z, photometric redshifts of galaxies)
using both simulated catalogs, and SDSS, DECam and Gaia photometry. We intend to make LSST-based value-added PhotoD catalogs publicly
available via the Rubin Science Platform with every LSST Data Release. 
\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
%% https://astrothesaurus.org
%% You will be asked to selected these concepts during the submission process
%% but this old "keyword" functionality is maintained in case authors want
%% to include these concepts in their preprints.
\keywords{Distance measure (395) --- Interstellar extinction (841) --- Photometry (1234) --- Stellar distance (1595) --- Two-color diagrams (1724)}

\include{intro}

\include{methodology}

\include{analysis}

\include{pipeline}

\include{discussion}

\begin{acknowledgements}
This work is financed within the Tenure Track Pilot Programme of the Croatian Science Foundation and the Ecole Polytechnique Fédérale
de Lausanne and the Project TTP-2018-07-1171 ``Mining the Variable Sky'', with the funds of the Croatian-Swiss Research Programme. 
Parts were based upon work supported by the National Science Foundation under Grant No. AST-2003196.

\v{Z}.I. acknowledges funding by the Fulbright Foundation and thanks the Ru{\dj}er Bo\v{s}kovi\'{c} Institute for hospitality. v{Z}.I, M.J. and N.\v{C}. acknowledge support from the DiRAC Institute in the Department of Astronomy at the University of Washington. B.A. acknowledges helpful discussions with Douglas Tucker and the members of the Rubin Observatory Stack Club.

The DiRAC Institute is supported through generous gifts from the Charles and
Lisa Simonyi Fund for Arts and Sciences and the Washington Research
Foundation. Funding for the SDSS and SDSS-II has been provided by the Alfred
P. Sloan Foundation, the Participating Institutions, the National
Science Foundation, the U.S. Department of Energy, the National
Aeronautics and Space Administration, the Japanese Monbukagakusho, the
Max Planck Society, and the Higher Education Funding Council for
England. The SDSS Web Site is \url{https://www.sdss.org/}.

This work has made use of data from the European Space Agency (ESA)
mission Gaia \url{https://www.cosmos.esa.int/gaia}, processed by the Gaia
Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/ dpac/consortium}). Funding for the
DPAC has been provided by national institutions, in particular, the
institutions participating in the Gaia Multilateral Agreement.
\end{acknowledgements}

\vspace{5mm}
\facilities{Sloan Digital Sky Survey, Gaia}
\software{numpy \citep{numpy}, matplotlib \citep{matplotlib}, seaborn \citep{seaborn}, scipy \citep{scipy}, pandas \cite{pandas}, 
                astropy \citep{astropy-1, astropy-2}, astroML \citep{2012cidu.conf...47V}, Jupyter \citep{jupyter}, Python \citep{python}.}

\appendix
%\include{appendix}

\bibliography{paper}{}
\bibliographystyle{aasjournal}

\end{document}